---
title: "False Positive Detection"
author: "Eric Matamoros"
date: '2022-10-21'
output:
  pdf_document: default
  html_document: default
---

## Importing modules

We first import the packages required from this RMarkdown.

```{r, message = FALSE}
library("RANN")
library("data.table")
library("doParallel")
library("ggplot2")
library("epifriends")
```

## Read mock data

Read mock data & specify parameters for running Epifriends.

```{r, message = FALSE}
setwd("./../") 
load("./data/mock_false_det.rda")
```

```{r, message = FALSE}
df <- mock_false_det
x <- df$x
y <- df$y
test <- df$test
positive = (test == 1)
link_d <- 0.02
min_neighbours <- 2
```

Check the distribution of positive and negative cases by plotting its distribution.

```{r, message = FALSE}
# Plot data
graph_base <- ggplot(df, aes(x = x, y = y, color = as.factor(test))) +
  geom_point(size = 3) +
  labs(title = "Distribution of Positive and Negative Cases")
print(graph_base)
```
In order to determine the number of false positives, the mean number of epifriends per cluster of positives will be computed with 500 different random infected cases distribution. These distributions are generated considering the global prevalence of the data to not impact the original infection rate.

Start by determining the prevalence of infection.

```{r, message = FALSE}
# prevalence infection
prevalence <- sum(test) / length(test)
print(paste0("The prevalence of infection is: ", prevalence))
```

Now we generate 500 random distributions with random infections with probability of infection equals to the prevalence computed previously.

```{r, message = FALSE}
# Generate random distributions
n_simulations = 100
scenarios <- generate_simulations(positions = df[,.(x,y)], test_result = test, 
                                 link_d = link_d, n_simulations = n_simulations,
                                 min_neighbours = min_neighbours)
head(scenarios, n = 20)
```

```{r, message = FALSE}
# Obtain mean epifriends per number of positives
random_dist <- scenarios[,.(mean_epi = mean(epifriends)), by = "num_pos"]
random_dist[, num_pos := as.numeric(num_pos)]
setorderv(random_dist, "num_pos")
# Plot data
graph_mean <- ggplot(random_dist, aes(x = num_pos, y = mean_epi,)) +
  geom_line() + geom_point(size = 3) +
  labs(title = "Mean number of epifriends per cluster of positives")
print(graph_mean)
```
The interpretation of the above chart is that you have a high probability of getting some epifriends of 2 isolated positives detected by pure randomness, while this probability decreases with increasing the number of isolated cases (since it's logically more complicated to get 4 positive cluster isolated than 2).

Now the catalogue with all the real data is computed, and the number of isolated cases as well.
```{r, message = FALSE}
cat <- catalogue(x = x, y = y, test, link_d, cluster_id = NULL, min_neighbours = min_neighbours)
cat_real <- data.table(
    'positive' = cat$epifriends_catalogue$positives,
    'negative' = cat$epifriends_catalogue$negatives
  )
  
cat_real <- cat_real[, .N, by = "positive"]
names(cat_real) <- c("num_pos", "epifriends")
cat_real[, num_pos := as.numeric(num_pos)]
head(cat_real)
```

Both data.frames are merged to compute the probability of a false detection. In this case there is 1 epifriend detected, while by randomness the mean epifriends of 2 positives is 18, meaning that it's likely the detected epifriends is a False Positive.

```{r, message = FALSE}
fp <- merge(cat_real, random_dist, by = c("num_pos"), how = "left")
fp[, prob_fd := mean_epi / epifriends]
head(fp)
```

Let's now visuallize the p-values and detected clusters of the epifriends, without considering the false detections.

```{r, message = FALSE}
plot1 <- scatter_pval(df[,.(x,y)], cat$cluster_id, positive, prevalence = NULL, cat$epifriends_catalogue)
plot1
```


Finally we can adjust the p-value based on the following formulas:


p-value = 1 - pbinom(npos - 1, ntotal, total_positives/total_n)
real-p-value = 1 - p-value

real-prob-false-detection =  1 - prob-df

adjusted-p-value = real-p-value * real-prob-false-detection
p-value = 1 - adjusted-p-value


Now let's recompute the catalogues but adjusting the p-value using the false detections.
```{r, message = FALSE}
catalogues <- catalogue(x = x, y =y, test, link_d, cluster_id = NULL,
                      min_neighbours = 2, max_p = 1, min_pos = 2, min_total = 2,
                      min_pr = 0, keep_null_tests = FALSE, in_latlon = FALSE, 
                      to_epsg = NULL, consider_fd = TRUE, n_simulations= 100,
                      verbose = FALSE)
```

```{r, message = FALSE}
plot2 <- scatter_pval(df[,.(x,y)], catalogues$cluster_id, positive, prevalence = NULL, catalogues$epifriends_catalogue, title = "P-value after removing False Detections")
plot2
```