---
title: "temporal_analysis"
author: "Eric Matamoros, Mikel Majewski, Arnau Pujol"
date: '2024-05-14'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing modules

We first import the packages required from this RMarkdown.

```{r, message = FALSE}
library("RANN")
library("data.table")
library("readxl")
library("chron")
library("ggplot2")
library("patchwork")
library("ggpubr")
library("repifriends")
```

## Create mock data 

The data created with the temporal_analysis.ipynb file of python and saved on this R project is readed:


```{r, message = FALSE}
setwd("./../") 
load("./data/datarand_temp.rda")
```

In order to work better with the data, some subdatasets are created:

```{r, message = FALSE}
mock_rand <- datarand_temp
mock_rand$date <- as.character(mock_rand$date)
x_rand <- mock_rand$x
y_rand <- mock_rand$y
test_rand <- mock_rand$test
positive_rand = (test_rand == 1)
date_rand <-mock_rand$date 
id_rand <- datarand_temp$id
```

The dates are changed to date format:

```{r, message = FALSE}
dtparts = t(as.data.frame(strsplit(date_rand," ")))
row.names(dtparts) = NULL
datesform <- chron(dates=dtparts[,1],times=dtparts[,2],format=c('y-m-d','h:m:s'))
dates <- as.numeric(datesform)
```

We also want to discretize the test variable, i.e. to convert it to integers

```{r, message = FALSE}
tests <- as.integer(test_rand)
```

In this plot we visualise the generated mock data. The colours of the points represent the date of the samples, and those marked with a red edge are the positive cases:

```{r, message = FALSE}
graph <- ggplot(mock_rand,aes(x=x, y=y))+
            geom_point(aes(fill = dates, colour = as.factor(test)), shape = 21, size = 2.5)+
            scale_fill_gradientn(colors = c("midnight blue", "turquoise", "yellow"))+
            scale_color_manual(values = c("grey", "red"), name = "Test", labels = c("Negative","Positive"))
graph
```

## Implementing EpiFRIenDs

Here we define the required parameters to run the method **temporal_catalogue** that outputs an EpiFRIenDs catalogue in each of the defined time steps:

```{r, message = FALSE}
#Parameters
#linking distance to link cases in space
link_d = 0.1
# Prevalence of the data in every time observation
prevalence =NULL
#Minimum number of positive cases within link_d to link them
min_neighbours = 2
#Width of the time window (in number of days) used to select the cases 
#in each time step. In this case, cases within 6 months are included
time_width = 180
#Number of days moved forward in each time step. In this case, each 
#time step include cases 3 months later than the previous step. Given 
#that time_width = 180, each step includes the last 90 from the previous 
#time step
time_steps = 90
#Initial date of first time step
min_date = min(datesform)
#Final date of last time step
max_date = max(datesform)
```

```{r, message = FALSE}
epi_catalogue_list <- temporal_catalogue(
  x = x_rand, y = y_rand, test_result = test_rand, dates =date_rand, link_d = link_d, prevalence = prevalence,
  use_link_d = TRUE, link_neighbours = NULL, min_neighbours=min_neighbours, time_width=time_width, 
  min_date=min_date, max_date=max_date, time_steps=time_steps, add_temporal_id = FALSE, dataset_ids = id_rand
)
```

```{r, message = FALSE}
epi_catalogue_list <- temporal_catalogue(
  x = x_rand, y = y_rand, test_result = test_rand, dates = date_rand, link_d = link_d,prevalence = prevalence,
  use_link_d = TRUE, link_neighbours = NULL, min_neighbours=min_neighbours, time_width=time_width,
  min_date=min_date, max_date=max_date, time_steps=time_steps, add_temporal_id = TRUE, linking_time = 3,
  linking_dist = 0.15, get_timelife = TRUE, dataset_ids = id_rand
)
```

Here we visualise the detected clusters in each step, showing one scatter plot for each time step with:

    * The cases not belonging to any cluster in grey (the positive cases marked with a red edge)

    * The cases belonging to a cluster colour-coded with their cluster ID (the positive cases marked with a red edge)

    * The p-value written next to each of the cluster

Notice that negative cases can in some cases belong to two (or more) different clusters (if they are closer than link_d to two (or more) positive cases from two (or more) different clusters. In this case, the colour has been assigned to the last of the plotted clusters.

```{r, message = FALSE}
step_num <- 0
while(min_date + time_steps*step_num + time_width <= max_date){
  print(paste0("Time step number", step_num)) 
  #select data in time window
  selected_data <- (dates >= min_date + time_steps*step_num)&
                    (dates <= min_date + time_steps*step_num + time_width)
  #Ploting all points with grey color and positive cases with red edge
  graph <- ggplot(mock_rand[test == 0,],aes(x=x, y=y))+
            geom_point(color = "grey", size = 2.5)+
            geom_point(data = mock_rand[test == 1,],aes(x=x, y=y), colour = "red", shape=21, fill = "grey", size = 2.5)+
            scale_color_manual(values = c("grey", "red"))
  #For each EpiFRIenDs cluster in this time step, plot its cases colour coded by the mean positivity rate
  # 
  for(epi in 1:length(epi_catalogue_list$temporal_catalogues[[step_num+1]]$id)){
    indeces <- epi_catalogue_list$temporal_catalogues[[step_num+1]]$indeces[[epi]]
    if(dim(mock_rand[selected_data,][indeces,])[1]>0){
      graph <- graph + geom_point(data = mock_rand[selected_data,][indeces,], aes(x = x, y = y, colour = as.factor(test)), shape = 21, size = 2.5, fill = as.integer(epi_catalogue_list$temporal_catalogues[[step_num+1]]$id[[epi]]))
    }
  }
  print(graph)
  
  step_num <- step_num + 1
  
}
```

Here we show the number of clusters detected as a function of time: 

```{r, message = FALSE}
num_clusters <- c()
for(i in epi_catalogue_list$temporal_catalogues){
  num_clusters <- append(num_clusters,length(i$id))
}

plot(x = epi_catalogue_list$mean_date, y = num_clusters, type = "l", xlab = "Dates", ylab = "Number of clusters")
```

Once we have the list of temporal **catalogues epi_catalogue_list**, we can add the temporal ID of the EpiFRIenDs clusters. This is done by assigning the same temporal ID to the clusters that are closer in time than the specificed linking time steps **linking_time** and closer in space that a specified **linking_dist**. Notice that **linking_dist** can be different than **link_d**, since the first is used to link clusters from different timesteps as assign to them the same temporal ID, and the latter is used to detect the clusters and assign their EpiFRIenDs ID.


```{r, message = FALSE}
#Number of time steps considered to link their temporal clusters
linking_time = 3
#Spatial distance of clusters within this linking_time to assign the same temporal ID
linking_dist = 0.15

#Get temporal IDs
epi_catalogue_list_id <- temporal_catalogue(
  x = x_rand, y = y_rand, test_result = test_rand, dates = date_rand, link_d = link_d, prevalence = prevalence,
  use_link_d = TRUE, link_neighbours = NULL, min_neighbours = min_neighbours, time_width = time_width, 
  min_date = min_date, max_date=max_date, time_steps=time_steps, add_temporal_id = TRUE, 
  linking_time =  linking_time, linking_dist = linking_dist, get_timelife = TRUE)
```






